---
title: "Tooth Growth Analysis"
author: "Martin Connolly"
date: "June 21, 2015"
output: pdf_document
---

# Exploratory Analysis

Loading data

```{r, echo=TRUE}
library(datasets)
library(ggplot2)
library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
test_data <- ToothGrowth
str(test_data)
summary(test_data)
```

Dose data are categorical rather than continuous, so need to convert this to a factor variable.

```{r, echo=TRUE}
test_data$dose <- factor(test_data$dose)
```


```{r, echo=TRUE, warning=FALSE}
ggplot(test_data, aes(x=dose, y=len)) +
        geom_violin(alpha=0.5) +
        geom_point(alpha=0.5) +
        labs(x="Dose",y="Length") + facet_grid(.~supp) +
        ggtitle("Dose versus length")

```

From the above chart, we can see that variability is not consistent, which we will need to take account of when performing any tests.

# Compare tooth length and supplement type

```{r, echo=TRUE}
len_OJ <- select(filter(test_data, supp == "OJ"), len)
len_VC <- select(filter(test_data, supp == "VC"), len)
test_supp <- t.test(len_OJ, len_VC, conf.level=.95, var.equal=FALSE, paired=FALSE, alternative ="two.sided")
test_supp
```


We can see that the p-value is `r test_supp$p.value` which is higher than 5% and the confidence interval contains zero, so we cannot reject the null hypothesis and we cannot say that the type of supplement alone will affect tooth growth.

# Compare tooth growth and dose

```{r, echo=TRUE}
#subset data based on dose
len_small_dose <- select(filter(test_data, dose == 0.5), len)
len_medium_dose <- select(filter(test_data, dose == 1), len)
len_large_dose <- select(filter(test_data, dose == 2), len)
```

## Compare effect on length of small dose (0.5) of supplement versus medium dose (1.0)

```{r, echo=TRUE}
test_dose_1 <- t.test(len_small_dose, len_medium_dose, conf.level=.95, var.equal=FALSE, paired=FALSE, alternative ="two.sided")
test_dose_1
```

## Compare effect on length of medium dose (1.0) of supplement versus large dose (2.0)

```{r, echo=TRUE}
test_dose_2 <- t.test(len_medium_dose, len_large_dose, conf.level=.95, var.equal=FALSE, paired=FALSE, alternative ="two.sided")
test_dose_2
```

## Compare effect on length of small dose (0.5) of supplement versus large dose (2.0)

```{r, echo=TRUE}
test_dose_3 <- t.test(len_small_dose, len_large_dose, conf.level=.95, var.equal=FALSE, paired=FALSE, alternative ="two.sided")
test_dose_3
```

# Conclusion

We have to reject the null hypothesis for all three t-tests and conclude that dosage of the supplements affect tooth growth.


